---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Reading the data:
```{r echo=TRUE}
unzip('activity.zip', exdir = "./data")
activities <- read.csv('./data/activity.csv', header=TRUE, sep = ',')

```

Process/transform the data (if necessary) into a format suitable for your analysis

Changing the data format from factor to data
```{r }
activities$date <- as.Date( activities$date, format = '%Y-%m-%d' )
```

We decide to sumarize the data by the number of steps by day
```{r echo=TRUE}
steps_by_day <- tapply(activities$steps, activities$date, function(x) sum(x,na.rm = TRUE))
steps_by_day <- as.data.frame(steps_by_day)
steps_by_day$steps_by_day <- as.numeric(steps_by_day$steps_by_day)

```

## What is mean total number of steps taken per day?

For this part of the assignment, we ignore the missing values in the dataset.

Making a histogram of the total number of steps taken each day:
```{r echo=TRUE}
library(ggplot2)

to.plot <- ggplot(steps_by_day,aes(x = steps_by_day)) + 
        geom_histogram() + 
        ggtitle("Histogram of total steps per day") +
        xlab("Number of steps made per day") +
        ylab("Frequency")
print(to.plot)
```
Calculate and report the mean and median total number of steps taken per day
```{r echo=TRUE}
mean(steps_by_day$steps_by_day, na.rm=TRUE)
median(steps_by_day$steps_by_day, na.rm=TRUE)
```

## What is the average daily activity pattern?

Making a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all days (y-axis)

```{r echo=TRUE}
steps_mean_by_day <- tapply(activities$steps, activities$interval, function(x) mean(x,na.rm = TRUE))
steps_mean_by_day <- as.data.frame(steps_mean_by_day)

plot(steps_mean_by_day$steps_mean_by_day, type="l", main="Average across all days based on intervals", xlab="Intervals", ylab="Steps Averaged")
```

Which 5-minute interval, on average across all the days in the dataset,
contains the maximum number of steps?
```{r echo=TRUE}
my_max = steps_mean_by_day[which.max(steps_mean_by_day$steps_mean_by_day),]
names(my_max)
paste("The interval with maximum number of steps is:", names(my_max))
```

## Imputing missing values

Calculating and reporting the total number of missing values in the dataset
(i.e. the total number of rows with NAs)
```{r echo=TRUE}
```

Filling in all of the missing values in the dataset. The strategy 
is to exchange the missing values with the mean values for intervals.
```{r echo=TRUE}
```

Creating a new dataset that is equal to the original dataset but with the
missing data filled in
```{r echo=TRUE}
```

Making a histogram of the total number of steps taken each day and Calculating
and reporting the mean and median total number of steps taken per day. 
Now we are able to awnser the following questions:
- Do these values differ from the estimates from the first part of the assignment?
- What is the impact of imputing missing data on the estimates of the total
daily number of steps?
```{r echo=TRUE}
```


## Are there differences in activity patterns between weekdays and weekends?







